### 2010-11-29 docs/tex/mimicry.tex
Modified with Dr. Grogono's comments

\begin{quote}
``It is not the strongest of the species that survives, nor the most intelligent that survives. It is the one that is the most adaptable to change. - Charles Darwin"
\end{quote}

\section{History}
Henry W. Bates first published in 1862 his findings about the similarities and dissimilarities between Heliconiinae and Ithomiinae butterflies, after 10 years of research in the Brazilian rain forest. For the next hundred years, it simulated heated discussion among all groups of people, scientists, philosophers, theologians, teachers and amateur naturalists. Bates collected ninety-four pieces of butterfly. He grouped them according to their similar appearance. He found butterflies having similar appearance, exhibiting morphological features which point to completely different species even families. Out of the ninety four species sixty seven are now classified as Ithomiinae, while twenty seven of them are Heliconiinae.

\section{Batesian Mimicry}
Even though Heliconiids are conspicuously colored, they are extremely abundant. They were also slow in mobility. Still predators in the surrounding area, mostly insectivorous birds do not prey on them, because of their inedible and unpalatable nature. Also because of this phenomenon other edible and palatable species such as ithomiinae and pieridae, pretend to be heliconiids and thus enjoy protection.

Repulsive animals, such as heliconiids are very conspicuously colored. Having this noticeable property, they are easily recalled by predators. Their wing pattern works as a warning to them. Once a predator has the knowledge of their inedible and unpalatable property, they would probably never attempt to try it again. As this is true, if any organism within close family and species, but being edible and having a deceptive resemblance to those conspicuously colored species will be avoided by the predators. Wickler \cite{wickler1986} expresses,
\begin{quote}
``Such unpalatable appearing and yet edible animals thus possess a false warning pattern, they 'act a part'. An actor is a mime, and so the representation of a false warning pattern was called \textit{mimicry}. Since Bates was the first to point out this phenomenon, it has received the name \textit{Batesian Mimicry} in his honor."
\end{quote}
In general, the animal which is avoided by predator for unpalatable behavior is called the \textbf{model} and the imitating animal is called the \textbf{mimic}.

\begin{figure}[H]
	\centering
	\includegraphics[scale=1]{images/Batesplate_ArM}
	\caption{Plate from Bates (1862) illustrating Batesian mimicry between Dismorphia species (top row, third row) and various Ithomiini (Nymphalidae) (second row, bottom row). \cite{bates1862}}
	\label{fig:batesian-butterfly}
\end{figure}

@@

\begin{quote}
``resemblance in external appearance, shapes and colors between members of widely distinct families."
\end{quote}

@@

\begin{quote}
``Mimicry is the close resemblance of one organism to another which, because it is unpalatable and conspicuous, is recognized and avoided by some predators at some times."
\end{quote}

@@

\begin{itemize}
	\item ``that the imitative species occur in the same area and occupy the same station as the imitated". This increases the probability that a predator will affect both parties concerned. Nevertheless, the model could live in Africa and the mimic in Europe, or vice versa, if the deceptive predator were a migratory bird.
	\item ``that the imitators are always the more defenseless". On the contrary, for the case of Mertensian mimicry among coral snakes the more offensive party may occasionally be the mimic.
	\item ``that the imitators are always less numerous in individuals". What is meant here is that the deceived animal should encounter the mimic less often than the model. This is true when the deceived animal must learn to recognize model and mimic and when positive and negative experiences carry equal weight. But it is known from various learning experiments that when an animal has learn about warning coloration, negative experiences or punishment stimuli (such as inedibility) can have a stronger and more lasting effect than positive experiences. Accordingly it should be expected that there would be an excess of mimics over models in proportion to the degree of predominance of negative experiences over positive onces. When the deceptive animal possesses an innate reaction to the model the number of mimics can be virtually unlimited. 
	\item ``that the imitators differ from the bulk of their allies". The most remarkable cases of mimicry are those where close relatives form divergent models. It is nevertheless highly possible that groups of closely related species may be similar mimics of the same model. Of course, it must always be remembered that species can look very different from their relatives because of adaptations which have nothing to do with mimicry. 
	\item ``that the imitation, however minute, is external and visible only, never extending to internal characters or to such as do not affect the external appearance." This criteria, like the previous one is aimed at defining mimicry as a special adaptation. But if one of two closely related species both of which are poisonous and possess similar warning coloration, should secondarily lose its poisonous qualities, it would become a mimic and yet resemble its model in almost all its non-mimetic characters.
\end{itemize}

@@

\section{Mullerian Mimicry}
Bates was not able to explain some phenomenon of mimicry. Occasionally two inedible unrelated butterfly species are amazingly similar in appearance. An explanation for this was provided by Fritz Muller in 1878. Like Bates, Muller observed and caught butterflies in Brazil. When there are multiple inedible species it is hard for predators to recognize each of them to know which one to consume and which one to avoid. Because of predator's limited memory, all these species still loose their number even after being inedible. So to save this loss, and to prevent more sacrifice of their own kind, inedible species from different family also tend to evolve to have similar appearance. This phenomenon is referred to as Mullerian Mimicry in the name of Fritz Muller.

According to Huheey \cite{huheey1988} ``Mullerian mimicry is normally viewed as a system of mutual warning coloration arising from the convergence of two or more warningly colored species, though two closely related species could also be Mullerian mimics through parallel evolution \cite{muller1879}." In Muller's view, predation load on each species are reduced when their warning signals are shared or converged. Naive young predators create the most impact in learning predation as they start their life with no knowledge or experience over the prey species. For cases like these, all unpalatable species should benefit from mutual advertising. If mutation gives opportunity for a single unpalatable prey to resemble a warningly colored species, it is expected to be more likely to survive predation than if the change was in opposite direction. Predators surviving on warningly colored species need to generalize; meaning if similar patterns signal unpalatability, predators recognition rate becomes much higher. Thus, variation in color pattern is tolerated, and mutations leading to even poor resemblance provides some protection.

It is worth observing the fact that Mullerian mimicry is not deception, as it is to the predator's advantage to be able to recognize all deterrent and noxious species. Preys resemblance does not have to be close enough to cause misidentification on the part of the predator, but merely similar enough to remind it of a bitter past experience with similar prey. As Huheey continues ``Thus although there is selection for convergence to perhaps the most common, average pattern, or more likely the one associated with the most deterrence (both from the intensity of the traumatic stimuli and the frequency of predators receiving them), the convergence need not be rapid or precise. Indeed, the resemblance among Mullerian mimics is often not close, and the existence of very close mimics raises doubts that the mimicry is Mullerian. "

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.75]{images/BatesMimButter}
	\caption{A very well-known examples of mimicry is the viceroy butterfly (top) which has a pattern very similar to the unpalatable monarch butterfly (bottom). It was considered for a long time that this is an ideal example of Batesian mimicry. But recently Brower discovered that the viceroy is actually just as noxious as the monarch, making this a case of Mullerian mimicry. \cite{brower1991} Image source: \href{http://en.wikipedia.org/wiki/Mullerian_mimicry}{Wikipedia}}
	\label{fig:mullerian-butterfly}
\end{figure}

@@

\paragraph{Punctuated Equilibrium}
In evolutionary biology, punctuated equilibrium is the theory which proposes that most sexually reproducing species will remain in an extended state called \textit{stasis} while experiencing little evolutionary change for most of their geological history. When evolution occurs it is localized in rare, rapid events of branching speciation, called cladogenesis. Cladogenesis is the process by which species split into two distinct species, rather than one species gradually transforming into another. Thus, ``punctuated equilibria is a model for discontinuous tempos of change (in) the process of speciation and the deployment of species in geological time" \cite{gould1977}.

@@

``According to punctuated equilibrium, therefore, evolution is not more of less both, gradual, and continual, as has been supposed for most purposes in most post-Darwinian theories. Rather, most morphological change is held to take place with in geologically short periods, separated by vastly longer periods of comparative stasis \cite{eldredge1972}. The rapid, large changes all occur during speciation, that is to say during the branching of the evolutionary tree (cladogenesis);" \cite{turner1988}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/Punctuated-equilibrium}
	\caption{Punctuated equilibrium, bottom, consists of morphological stability and rare bursts of evolutionary change, while the top explains phyletic gradualism. Image source: \href{http://en.wikipedia.org/wiki/Punctuated_equilibrium}{Wikipedia}}
	\label{fig:punctuated-equilibrium}
\end{figure}

>>>>

\begin{quote}
\textsl{``It is not the strongest of the species that survives, nor the most intelligent that survives. It is the one that is the most adaptable to change." - Charles Darwin}
\end{quote}

\section{History}
Henry W. Bates first published in 1862 his findings about the similarities and dissimilarities between Heliconiinae and Ithomiinae butterflies, after 10 years of research in the Brazilian rain forest. For the next hundred years, it simulated heated discussion among all groups of people, scientists, philosophers, theologians, teachers and amateur naturalists. Bates collected ninety-four pieces of butterfly. He grouped them according to their similar appearance. He found butterflies having similar appearance, exhibiting morphological features which point to completely different species even families. Out of the ninety four species, sixty seven are now classified as Ithomiinae, while twenty seven of them are Heliconiinae.

\section{Batesian Mimicry}
Even though Heliconiids are conspicuously colored, they are extremely abundant. They are also slow in mobility. Still predators in the surrounding area, mostly insectivorous birds do not prey on them, because of their inedible and unpalatable nature. Also because of this phenomenon other edible and palatable species such as ithomiinae and pieridae, pretend to be heliconiids and thus enjoy protection.

Repulsive animals, such as heliconiids are very conspicuously colored. Having this noticeable property, they are easily recalled by predators. Their wing pattern works as a warning to predators. Once a predator has the knowledge of their inedible and unpalatable property, they would probably never attempt to try it again. As this is true, if any organism within close family and species, but being edible and having a deceptive resemblance to those conspicuously colored species will be avoided by the predators. Wickler \cite{wickler1986} expresses,
\begin{quote}
\textsl{``Such unpalatable appearing and yet edible animals thus possess a false warning pattern, they 'act a part'. An actor is a mime, and so the representation of a false warning pattern was called \textit{mimicry}. Since Bates was the first to point out this phenomenon, it has received the name \textit{Batesian Mimicry} in his honor."}
\end{quote}
In general, the animal which is avoided by predator for unpalatable behavior is called the \textbf{model} and the imitating animal is called the \textbf{mimic}.

\begin{figure}[H]
	\centering
	\includegraphics[scale=1]{images/Batesplate_ArM}
	\caption[Plate from Bates (1862) illustrating Batesian mimicry]{Plate from Bates (1862) illustrating Batesian mimicry between Dismorphia species (top row, third row) and various Ithomiini (Nymphalidae) (second row, bottom row). \cite{bates1862}}
	\label{fig:batesian-butterfly}
\end{figure}

@@

\begin{quote}
\textsl{``resemblance in external appearance, shapes and colors between members of widely distinct families."}
\end{quote}

@@

\begin{quote}
\textsl{``Mimicry is the close resemblance of one organism to another which, because it is unpalatable and conspicuous, is recognized and avoided by some predators at some times."}
\end{quote}

@@

\begin{itemize}
	\item \textsl{``that the imitative species occur in the same area and occupy the same station as the imitated".} This increases the probability that a predator will affect both parties concerned. Nevertheless, the model could live in Africa and the mimic in Europe, or vice versa, if the deceptive predator were a migratory bird.
	\item \textsl{``that the imitators are always the more defenseless".} On the contrary, for the case of Mertensian mimicry among coral snakes the more offensive party may occasionally be the mimic.
	\item \textsl{``that the imitators are always less numerous in individuals".} What is meant here is that the deceived animal should encounter the mimic less often than the model. This is true when the deceived animal must learn to recognize model and mimic and when positive and negative experiences carry equal weight. But it is known from various learning experiments that when an animal has learned about warning coloration, negative experiences or punishment stimuli (such as inedibility) can have a stronger and more lasting effect than positive experiences. Accordingly it should be expected that there would be an excess of mimics over models in proportion to the degree of predominance of negative experiences over positive onces. When the deceptive animal possesses an innate reaction to the model the number of mimics can be virtually unlimited. 
	\item \textsl{``that the imitators differ from the bulk of their allies".} The most remarkable cases of mimicry are those where close relatives form divergent models. It is nevertheless highly possible that groups of closely related species may be similar mimics of the same model. Of course, it must always be remembered that species can look very different from their relatives because of adaptations which have nothing to do with mimicry. 
	\item \textsl{``that the imitation, however minute, is external and visible only, never extending to internal characters or to such as do not affect the external appearance."} This criteria, like the previous one is aimed at defining mimicry as a special adaptation. But if one of two closely related species both of which are poisonous and possess similar warning coloration, should secondarily lose its poisonous qualities, it would become a mimic and yet resemble its model in almost all its non-mimetic characters.
\end{itemize}

@@

\section{Mullerian Mimicry}
Bates was not able to explain some phenomena of mimicry. Occasionally two inedible unrelated butterfly species are amazingly similar in appearance. An explanation for this was provided by Fritz Muller in 1878. Like Bates, Muller observed and caught butterflies in Brazil. When there are multiple inedible species it is hard for predators to recognize each of them to know which one to consume and which one to avoid. Because of predator's limited memory, all these species still lose their number even after being inedible. So to save this loss, and to prevent more sacrifice of their own kind, inedible species from different family also tend to evolve to have similar appearance. This phenomena is referred to as Mullerian Mimicry in the name of Fritz Muller.

According to Huheey \cite{huheey1988} \textsl{``Mullerian mimicry is normally viewed as a system of mutual warning coloration arising from the convergence of two or more warningly colored species, though two closely related species could also be Mullerian mimics through parallel evolution" \cite{muller1879}.} In Muller's view, predation load on each species are reduced when their warning signals are shared or converged. Naive young predators create the most impact in learning predation as they start their life with no knowledge or experience over the prey species. For cases like these, all unpalatable species should benefit from mutual advertising. If mutation gives opportunity for a single unpalatable prey to resemble a warningly colored species, it is expected to be more likely to survive predation than if the change was in opposite direction. Predators surviving on warningly colored species need to generalize; meaning if similar patterns signal unpalatability, predators recognition rate becomes much higher. Thus, variation in color pattern is tolerated, and mutations leading to even poor resemblance provides some protection.

It is worth observing the fact that Mullerian mimicry is not deception, as it is to the predator's advantage to be able to recognize all deterrent and noxious species. Preys resemblance does not have to be close enough to cause misidentification on the part of the predator, but merely similar enough to remind it of a bitter past experience with similar prey. As Huheey continues \textsl{``Thus although there is selection for convergence to perhaps the most common, average pattern, or more likely the one associated with the most deterrence (both from the intensity of the traumatic stimuli and the frequency of predators receiving them), the convergence need not be rapid or precise. Indeed, the resemblance among Mullerian mimics is often not close, and the existence of very close mimics raises doubts that the mimicry is Mullerian."}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.75]{images/BatesMimButter}
	\caption[A very well-known example of mimicry, the viceroy butterfly]{A very well-known example of mimicry is the viceroy butterfly (top) which has a pattern very similar to the unpalatable monarch butterfly (bottom). It was considered for a long time that this is an ideal example of Batesian mimicry. But recently Brower discovered that the viceroy is actually just as noxious as the monarch, making this a case of Mullerian mimicry. \cite{brower1991} Image source: \href{http://en.wikipedia.org/wiki/Mullerian_mimicry}{Wikipedia}}
	\label{fig:mullerian-butterfly}
\end{figure}

@@

\paragraph{Punctuated Equilibrium}
In evolutionary biology, punctuated equilibrium is the theory which proposes that most sexually reproducing species will remain in an extended state called \textit{stasis} while experiencing little evolutionary change for most of their geological history. When evolution occurs it is localized in rare, rapid events of branching speciation, called cladogenesis. Cladogenesis is the process by which species split into two distinct species, rather than one species gradually transforming into another. Thus, \textsl{``punctuated equilibria is a model for discontinuous tempos of change (in) the process of speciation and the deployment of species in geological time" \cite{gould1977}}.

@@

\textsl{``According to punctuated equilibrium, therefore, evolution is not more of less both gradual and continual, as has been supposed for most purposes in most post-Darwinian theories. Rather, most morphological change is held to take place within geologically short periods, separated by vastly longer periods of comparative stasis \cite{eldredge1972}. The rapid, large changes all occur during speciation, that is to say during the branching of the evolutionary tree (cladogenesis);" \cite{turner1988}}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/Punctuated-equilibrium}
	\caption[Punctuated equilibrium vs. phyletic gradualism]{Punctuated equilibrium, bottom, consists of morphological stability and rare bursts of evolutionary change, while the top explains phyletic gradualism. Image source: \href{http://en.wikipedia.org/wiki/Punctuated_equilibrium}{Wikipedia}}
	\label{fig:punctuated-equilibrium}
\end{figure}

### 2010-11-29 docs/tex/model.tex
Changed the name of the chapters.
Some updates. Nothing significant.
Modified with Dr. Grogono's comments

%Chapter describing the model
\chapter{Modeling the evolution of mimicry}

\begin{quotation}
``Modeling, it should be clear is an art form. It depends on the experience and taste of the modeler. In this it is much like cartooning. The modeler (cartoonist) must decide which features to make salient (exaggerate), and which features to eliminate (avoid), in order to answer the questions (make the political point)" - John H. Holland \cite{holland1996}.
\end{quotation}

@@

The environment is designed as three dimensional, while the space will be of toroidal nature. This idea has been taken from the Laws and Life project by Peter Grogono \cite{grogono2003}.

@@

%Explain the results for the second model by Frank and Noble
\paragraph{Results}
According to the first experiment, which started with 20 unpalatable prey species, hypothesis 1 of a single large mimicry ring was not established. Their was existence of multiple mimicry rings of very different frequency. Second experiment was carried out with some palatable species along with the unpalatable once which was able to reach conditions of Batesian mimicry also with less mimicry rings, were hypothesis 2 borne out. It can be observed that,

\begin{quote}
\textsl{The presence of Batesian mimics would provide positive selection pressure on mutants and would, therefore, increase the probability that they would evolve an initial resemblance to another unpalatable species. Also, Batesian pressure on mimicry rings has the potential to push one ring into the range of another, helping to bridge a large phenotypic difference between them. \cite{franks2003}}
\end{quote}

\section{FormAL Framework}
The ``FormAL framework" is a collection of ideas and concepts taken from Peter Grogono's FormAL(Formal Artificial Life) project \cite{grogono2003} and are used to build a framework for modeling the evolution of mimicry. The primary goal of the FormAL project was ``\textit{To study the emergence of complexity}". While the principal behind it was ``\textit{not to include a variable in an agent unless the variable is genetically controlled (or, at least, genetically influenced)}". We will see that this principal was not completely established in simulating mimicry as it was not possible to come up with a genetic representation of Hopfield Network which was used as a pattern recognizer for the predator species.

@@

\begin{table}[H]
\centering
\begin{tabular}{cc}
		\multicolumn{2}{c}{ \includegraphics[scale=0.50]{images/cells-front} } \\
		\includegraphics[scale=0.40]{images/cells-top} & \includegraphics[scale=0.40]{images/cells-side} \\
\end{tabular}
\caption{Three dimensional representation of the environment divided in cells. Presence of different species of agents inside.}
\label{tab:3-d-environment-images}
\end{table}

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=5]{images/CA_rule30s}
	\caption{Cellular Automata Rule 30.
	Image source: \href{http://en.wikipedia.org/wiki/Cellular_automata}{Wikipedia}}
	\label{fig:cellular-automata-rule-30}
\end{figure}

\begin{table}[H]
	\centering
	\begin{tabular}{| l | c | c | c | c | c | c | c | c |}
	  \hline
	  Current Pattern & 111 & 110 & 101 & 100 & 011 & 010 & 001 & 000 \\ \hline
	  New state of center cell & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 0 \\
	  \hline
	\end{tabular}
	\caption{Cellular Automata Rule}
	\label{tab:cellular-automata-rule}
\end{table}

@@

\subsection{Genome}
The Genome of the prey species consists of 17 bits as presented in Table \ref{tab:prey-genome}. The first eight bit is the rule, which is used to generate Cellular Automata pattern. The next two bits are used to represent palatability of the organism explained detail in Section \ref{sec:genetic-palatability-representation}. The next six bits is the magnitude of the force with which mobility of the organism is calculated. The 17th bit is used to evaluate reproduction capability of the organism, depending on which the prey species is either capable or not capable to participate in reproduction.

@@

\subsection{Genetic representation of palatability}
\label{sec:genetic-palatability-representation}
The palatability of each prey species is fixed and has been represented with 2 bit of the genome giving it a range of 0 to 3 with four levels of palatability. The combinations are as follows:

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|}
		\hline
			\textbf{Gene (Index 8 to 9)} &	\textbf{Palatable} \\ \hline
			00									& True 			\\ \hline
			01									& True 			\\ \hline
			10									& False 		\\ \hline
			11									& False 		\\
		\hline
	\end{tabular}
	\caption{Genetic representation of Palatability}
	\label{tab:genetic-representation-palatability}
\end{table}

@@

\subsubsection{Reproduction}
Every prey species starts reproducing when it reaches the ``Reproductive age limit". If it is capable of reproducing, which is decided based on its 17th bit gene, the prey will randomly select another prey species with similar pattern and palatability from the same cell and mate with it. Now the other prey also needs to be genetically capable of reproduction and needs to reach its age limit. A new genome is created from the existing genome of the two prey by applying single point crossover operation. Mutation is performed separately on the pattern gene and the rest of the genome, with two different rates to control them using the values in Table \ref{tab:prey-control-parameters}. So there is two point mutation for the genome.

@@

Predators in the system are designed to provide selection pressure to \textit{models} and \textit{mimics} for the evolution of mimicry. Similar to prey species, they are agents in the FormAL environment capable of mobility and reproduction. In addition to it, these agents are equipped with Hopfield Network Memory to be able to learn and recognize patterns of the prey species. Their mobility and reproduction capability is controlled at the genetic level, while their memory is not genetically controlled, as it was not possible to come up with a genetic representation for Hopfield Network. So every new predator species gets to born with zero memory and with no inharitance from parents. A set of parameters are defined to control predators' population and learning ability in the environment. Details of these parameters are given in Table \ref{tab:predator-control-parameters}.

@@

\begin{quote}
\textsl{When an axon of cell A is near enough to excite a cell B and repeatedly or persistently takes part in firing it, some growth process or metabolic changes take place in one or both cells such that A's efficiency as one of the cells firing B is increased.}
\end{quote}

@@

In this way the predator memory is limited to ``Hopfield Maximum Memory Size". After reaching this memory predator will not store any more new pattern but will try to associate with the existing ones it has already stored. Algorithm \ref{algo:algorithm-attack-prey} and \ref{algo:algorithm-attemptToKill} explains the process in every detail.

@@

\subsection{Genome}
Each predators has a 5 bit genome as explained in Table \ref{tab:predator-genome}. The first four bit represents its mobility and the magnitude of the force component with which it moves around the environment. The last bit controls reproduction capability of the specific species.

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.40]{images/predators-40}
	\caption{A population of 40 predators distributed over the environment.}
	\label{fig:predators-40}
\end{figure}

@@

There are two control parameters that effect the reproduction of predator species. The first is "Reproduction Age Limit". This is the minimum age a predator has to reach before starting to involve itself for reproduction. This parameter has been set to 500 iterations during the result and analysis section of the thesis. The second parameter is "Reproduction Interval" which has been varied in the simulation from 1000 to 3000 iteration depending on the population of palatable prey species. This is a very important parameter for the simulation as it determines the overall predator population and its rate of increment. Depending on this value we can control the rate of predation on prey species, which on the other hand controls the rate of mimetic behavior of the overall prey population.

>>>>

%Chapter describing the model
\chapter{The Model: Evolution of Mimicry}

\begin{quote}
\textsl{``Modeling, it should be clear is an art form. It depends on the experience and taste of the modeler. In this it is much like cartooning. The modeler (cartoonist) must decide which features to make salient (exaggerate), and which features to eliminate (avoid), in order to answer the questions (make the political point)" - John H. Holland \cite{holland1996}.}
\end{quote}

@@

The environment is designed as three dimensional, while the space will be of toroidal nature. This idea has been taken from the \textsl{Laws and Life} project by Peter Grogono \cite{grogono2003}.

@@

%Explain the results for the second model by Frank and Noble
\paragraph{Results}
According to the first experiment, which started with 20 unpalatable prey species, hypothesis 1 of a single large mimicry ring was not established. There was existence of multiple mimicry rings of very different frequency (population). Second experiment was carried out with some palatable species along with the unpalatable once which was able to reach conditions of Batesian mimicry also with less mimicry rings, where hypothesis 2 borne out. It can be observed that,

\begin{quote}
\textsl{``The presence of Batesian mimics would provide positive selection pressure on mutants and would, therefore, increase the probability that they would evolve an initial resemblance to another unpalatable species. Also, Batesian pressure on mimicry rings has the potential to push one ring into the range of another, helping to bridge a large phenotypic difference between them." \cite{franks2003}}
\end{quote}

\section{FormAL Framework}
The \textsl{``FormAL framework"} is a collection of ideas and concepts taken from Peter Grogono's FormAL(Formal Artificial Life) project \cite{grogono2003} and are used to build a framework for modeling the evolution of mimicry. The primary goal of the FormAL project was ``\textit{To study the emergence of complexity}". While the principal behind it was ``\textit{not to include a variable in an agent unless the variable is genetically controlled (or, at least, genetically influenced)}". We will see that this principal was not completely established in simulating mimicry as it was not possible to come up with a genetic representation of Hopfield Network which was used as a pattern recognizer for the predator species.

@@

\begin{table}[H]
\centering
\begin{tabular}{cc}
		\multicolumn{2}{c}{ \includegraphics[scale=0.60]{images/cells-front} } \\
		\includegraphics[scale=0.50]{images/cells-top} & \includegraphics[scale=0.50]{images/cells-side} \\
\end{tabular}
\caption[Three dimensional representation of the environment]{Three dimensional representation of the environment divided in cells. Presence of different species of agents inside.}
\label{tab:3-d-environment-images}
\end{table}

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=5]{images/CA_rule30s}
	\caption[Cellular Automata]{Cellular Automata Rule 30.
	Image source: \href{http://en.wikipedia.org/wiki/Cellular_automata}{Wikipedia}}
	\label{fig:cellular-automata-rule-30}
\end{figure}

\begin{table}[H]
	\centering
	\begin{tabular}{| l | c | c | c | c | c | c | c | c |}
	  \hline
	  Current Pattern & 111 & 110 & 101 & 100 & 011 & 010 & 001 & 000 \\ \hline
	  New state of center cell & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 0 \\
	  \hline
	\end{tabular}
	\caption{Cellular Automata rule}
	\label{tab:cellular-automata-rule}
\end{table}

@@

\subsection{Genome}
The Genome of the prey species consists of 17 bits as presented in Table \ref{tab:prey-genome}. The first eight bits is the rule, which is used to generate Cellular Automata pattern. The next two bits are used to represent palatability of the organism explained detail in Section \ref{sec:genetic-palatability-representation}. The next six bits is the magnitude of the force with which mobility of the organism is calculated. The 17th bit is used to evaluate reproduction capability of the organism, depending on which the prey species is either capable or not capable to participate in reproduction.

@@

\subsection{Genetic representation of palatability}
\label{sec:genetic-palatability-representation}
The palatability of each prey species is fixed and has been represented with 2 bits of the genome giving it a range of 0 to 3 with four levels of palatability. The combinations are as follows:

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|}
		\hline
			\textbf{Gene (Index 8 to 9)} &	\textbf{Palatable} \\ \hline
			00									& True 			\\ \hline
			01									& True 			\\ \hline
			10									& False 		\\ \hline
			11									& False 		\\
		\hline
	\end{tabular}
	\caption{Genetic representation of palatability}
	\label{tab:genetic-representation-palatability}
\end{table}

@@

\subsubsection{Reproduction}
Every prey species starts reproducing when it reaches the \textsl{``Reproductive age limit"}. If it is capable of reproducing, which is decided based on its 17th bit gene, the prey will randomly select another prey species with similar pattern and palatability from the same cell and mate with it. Now the other prey also needs to be genetically capable of reproduction and needs to reach its age limit. A new genome is created from the existing genome of the two prey by applying single point crossover operation. Mutation is performed separately on the pattern gene and the rest of the genome, with two different rates to control them using the values in Table \ref{tab:prey-control-parameters}. So there is two point mutation for the genome.

@@

Predators in the system are designed to provide selection pressure to \textit{models} and \textit{mimics} for the evolution of mimicry. Similar to prey species, they are agents in the FormAL environment capable of mobility and reproduction. In addition to it, these agents are equipped with Hopfield Network Memory to be able to learn and recognize patterns of the prey species. Their mobility and reproduction capability is controlled at the genetic level, while their memory is not genetically controlled, as it was not possible to come up with a genetic representation for Hopfield Network. So every new predator species gets to born with zero memory and with no inheritance from parents. A set of parameters are defined to control predators' population and learning ability in the environment. Details of these parameters are given in Table \ref{tab:predator-control-parameters}.

@@

\begin{quote}
\textsl{``When an axon of cell A is near enough to excite a cell B and repeatedly or persistently takes part in firing it, some growth process or metabolic changes take place in one or both cells such that A's efficiency as one of the cells firing B is increased."}
\end{quote}

@@

In this way the predator memory is limited to \textsl{``Hopfield Maximum Memory Size"}. After reaching this memory predator will not store any more new pattern but will try to associate with the existing ones it has already stored. Algorithm \ref{algo:algorithm-attack-prey} and \ref{algo:algorithm-attemptToKill} explains the process in every detail.

@@

\subsection{Genome}
Each predator has a 5 bit genome as explained in Table \ref{tab:predator-genome}. The first 4 bits represent its mobility and magnitude of the force component with which it moves around the environment. The last bit controls reproduction capability of each species.

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.60]{images/predators-40}
	\caption[A population of predators]{A population of 40 predators distributed over the environment.}
	\label{fig:predators-40}
\end{figure}

@@

There are two control parameters that effect the reproduction of predator species. The first is \textsl{"Reproduction Age Limit"}. This is the minimum age a predator has to reach before starting to involve itself for reproduction. This parameter has been set to 500 iterations during the result and analysis section of the thesis. The second parameter is \textsl{"Reproduction Interval"} which has been varied in the simulation from 1000 to 3000 iteration depending on the population of palatable prey species. This is a very important parameter for the simulation as it determines the overall predator population and its rate of increment. Depending on this value we can control the rate of predation on prey species, which on the other hand controls the rate of mimetic behavior of the overall prey population.
